---
title: "APIs in R"
author: "Team Hanley NLP Working Group"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
---

## What is an API?
An API stands for Application Program Interface, which may sound complicated, but all it is is a way for different programs to talk to each other. In this tutorial, we will go over a particular subset of APIs called RESTful APIs, which is simply a design pattern or type of APIs that is designed to be easy to use.

But before we go over RESTful APIs in particular, we need to define what makes an API an API. Prepare yourself for some jargon. 

- **Client**: a client is the program or person who is using the API. So if you made a shiny app that used an API, than your shiny app would be the client. If you built a script than you or that script would be the client. 
- **Resource**: this is the information you are trying to pull back with your API. In our example below, you will be pulling back information on Pokemon, so the pokemon is the resource. Keep in mind that each resource will have a unique identifier. 

So using the technical language we have learned above, we can think of an API as a way for a **client** to pull **resources** from a server.

### How Can We Interact with an API?

When we use an API to interact with a resource, we are making a **request** to a server. [There are many different types of requests](https://developer.mozilla.org/en-US/docs/Web/HTTP/Methods) that we could make, but we will go through the two most common here. 

- `GET`: This request is for **getting** information from a server about a resource. As far as data analysis goes, this is likely the only way we will be interacting with a server
- `POST`: This request is for **posting** information to a specified resource. For example, you could post a tweet using an API. This is how twitter bots operate. 

Let's do our first `GET` request now, but using our browser as a client instead of code. Click this [link](https://pokeapi.co/api/v2/pokemon/ditto). You should see a bunch of unintelligible text. This is response from our `GET` request. What's more important (and hopefully a bit demystifying) is the url. Let's take a closer look:

`https://pokeapi.co/api/v2/pokemon/ditto`

What we pulled back was some information about the pokemon `ditto`, and as you can see, the last part of that url is the word `ditto`. What if we were to change this to `pikachu`. In that case, we would get information back about `pikachu`. 

The best part of APIs is that they use a common design pattern to make different requests. That way requesting on pikachu versus ditto only means changing one thing. 

### What is a RESTful API?

REST stands for **RE**presentional **S**tate **T**ransfer. Make sense? No? We can break that down further. 

The state of resources on the web is constantly changing. How many followers a twitter user has, for example, changes all the time. What a RESTful API does then is provide a *representation* of the *state* of a particular resource at any given moment. If the resource is a twitter user, than the state of the resource could be things like its follower count, number of tweets, or favorited tweets.


## Querying an APIs in R

### jsonlite

APIs oftentimes return data in the form of JSON files, which can be thought of as akin to lists in R. When this is the case (and when life is easy and you don't need to authenicate), you can use the `jsonlite` to directly access the results. 

```{r, message = F, warning = F}

library(jsonlite)

# this will return a list with information about ditto
ditto = fromJSON("https://pokeapi.co/api/v2/pokemon/ditto")

ditto$species$name

```

## Problem 1: Building a Pokemon Dataset {.tabset .tabset-fade}

### Problem
Using the pokemon API from above, build a data frame that includes the name, height, weight, base experiance, and primary type for at least four pokemon. 

**Hints**

- Use `fromJSON()` to pull back 
- Use a `for loop` to loop through your chosen pokemon
- `paste0()` can be used to change the end point of your API

If you are getting stuck, I would recommend pulling back just a single pokemon and then looking for the information that you need. 

### Solution

There will definitely be many different ways to accomplish this but I will use a combination of a `for loop`, `append()`, and `paste0()`

```{r, message = F, warning = F}

library(jsonlite)
library(kableExtra)
library(dplyr)

pokemon = c("pikachu", "ditto", "eevee", "charizard")

# initialize outside of for loop
names = c()
height = c()
weight = c()
base_experience = c()
primary_type = c()

# loop through different pokemon
for(i in 1:length(pokemon)){
  
  # this will return a list with information about ditto
  data = fromJSON(paste0("https://pokeapi.co/api/v2/pokemon/", pokemon[i]))
  
  # append will add the result to the end of the vector
  names = append(names, data$species$name)
  height = append(height, data$height)
  weight = append(weight, data$weight)
  base_experience = append(base_experience, data$base_experience)
  primary_type = append(primary_type, data$types$type$name[1])
}

# data will compile the vectors into a dataframe
pokemon_df = data.frame(
  names,
  height,
  weight,
  base_experience,
  primary_type
)

kable(head(pokemon_df), format = "html")%>%
  kable_styling("striped")

```


## Problem 2: Make a Movie Dataset {.tabset .tabset-fade}

### Problem

[The Open Movie Database](http://www.omdbapi.com/apikey.aspx) has an API that allows users to pull data on up to 1,000 films a day for free. After you receive the API key in an email, activate the key following the link in the email. Then make a data frame that contains the title, year, MPAA rating, runtime, genre, and actors for a few of your favorite movies!

<br>
<br>
<br>

### Solution

```{r, message = F, warning = F}


library(jsonlite)
library(kableExtra)
library(dplyr)

films = c("Batman+Begins", "The+Incredibles", "The+Godfather", "Parasite")

 titles = c()
 years = c()
 ratings = c()
 runtimes = c()
 genres = c()
 actors = c()
 
 
for(i in 1:length(films)){

  data = fromJSON(paste0("http://omdbapi.com/?apikey=765ce2a6&t=", films[i]))
  
  titles = append(titles, data$Title)
  years = append(years, data$Year)
  ratings = append(ratings, data$Rated)
  runtimes = append(runtimes, data$Runtime)
  genres = append(genres, data$Genre)
  actors = append(actors, data$Actors)
  }

 film_df = data.frame(
   titles,
   years,
   ratings,
   runtimes,
   genres,
   actors
)

kable(head(film_df), format = "html")%>%
  kable_styling("striped")

```
