---
title: "Exploring Text Data"
author: "Team Hanley NLP Working Group"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary Statistics with Words

The first step in any data science project should always be to get a sense of your data. For NLP, a lot of exploratory data analysis revolves around counting the frequencies of different terms and plotting it in different ways. This can be as simple as a bar chart looking at the count of distinct words, to a word clouds, to something as complex as a TF-IDF. 

Before we go any further, we first need to set up our workspace. We will primarily be working with `tidytext` to keep ourselves in the `tidyverse` and will be pulling data from `gutenbergr`, which is a repository of free classic texts. Let's start by loading in our data, Metamorphasis by Kafka, and unnesting it into tokens.

```{r warning=FALSE, message=FALSE}
library(gutenbergr)
library(wordcloud)
library(ggplot2)
library(tidytext)
library(dplyr)

meta = gutenberg_download("5200") 

tokens = meta %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) 

```

Next we can create a simple barplot looking at the top ten words (not including stop words of course) used in the book. 

```{r warning=FALSE, message=FALSE}

word_count = tokens%>%
  group_by(word)%>%
  summarise(count = n())%>%
  arrange(desc(count))%>%
  slice(1:10)

ggplot(data = word_count)+
  geom_bar(aes(x = word, y = count), stat = "identity",  fill = "#6699cc")+
  theme_classic()

```

### Word Clouds with wordcloud

Another way to present summary statistics is with a word cloud. For the unintiated, a word cloud is just a collection of words with the size of the word determined by its prevelance. The natural disadvantage of this being that longer words will appear larger than smaller words, thus making them look more prevelant.  

Actually creating the wordcloud is very easy. Just wrap a character vector in `wordcloud()`. We can remove some of the less frequent words using either the `min.freq` argument or the `max.words` argument. 

```{r warning=FALSE, message=FALSE}

wordcloud(tokens$word, max.words = 75, colors=brewer.pal(6, "Dark2"))

```

## TF-IDF

So far, we have 

TF-IDF, or Term Frequency and Inverse Document Frequency is a way to numerically give importance to a word or phrase in a given text document relative to a collection of other text documents. 

In this example, we will examine how important a word is in a chapter of Metamorphosis in comparison to the entire novel. Our dataframe 'meta' contains the text from Metamorphosis, however there is no column designating chapters in the novel. By looking through the meta dataframe, we can see that there are chapter divides at rows 640 and 1296. Let's start by adding a chapter column to our dataframe, dividing Metamorphosis into it's 3 chapters.


```{r warning=FALSE, message=FALSE}

meta$chapter = NA
meta$chapter[1:639] = 1
meta$chapter[640:1295] = 2
meta$chapter[1296:nrow(meta)] = 3

meta <- meta[,2:3]
```

Next, we need to create a dataframe that breaks down our text into one word per row using unnest_tokens(). Here, we can count how many times words occur in each chapter. This is our term frequency (tf).

```{r warning=FALSE, message=FALSE}
book_words <- meta%>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  add_count(chapter, word, sort = TRUE)

book_words <- distinct(book_words)

book_words %>%
  arrange(desc(n))
```

Term frequency alone can tell us which words or phrases occur the most in a given document or collection of documents. This is helpful to some extent, however some of the words or phrases with the highest term frequency may not be that important, or rather they may not give us much insight into what the document or collection of documents is about. In this example, we can see that "the" is the most frequent term in each chapter,giving us no insight into the contents of these chapters. 

The IDF (inverse document frequency) of a word in a collection of documents can be understood as:

idf(word) = ln(total number of documents/number of documents containing word)

(ln is the natural log)

In this example, the number of documents is 3 since we have 3 total chapters. A term with an idf score of 0 is a term that occurs very frequently across the entire collection of text. The word "the" occurs in all 3 chapters. 

IDF example: 

idf("the") = ln(3/3) = 0

The bind_tf_idf() function gets the tf, idf, and tf-idf scores for each word in our dataset.

```{r warning=FALSE, message=FALSE}
book_words.2 <- book_words %>%
  bind_tf_idf(word,chapter, n)

book_words.2

book_words.2 %>%
  arrange(desc(tf_idf)) 

book_words.2 %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(chapter) %>% 
  top_n(15) %>% 
  ungroup() %>%
  ggplot(aes(word, tf_idf, fill = chapter)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  facet_wrap(~chapter, ncol = 2, scales = "free") +
  coord_flip()


```

Looking at the words with the highest tf-idf scores in each chapter, we can see which words are more important to individual chapters. Gregor Samsa is one of the main characters in this novel. The term "Gregor" occurred too frequenty throughout the entire book, however his last name, "Samsa" was probably not used as often since we can see it has the highest tf-idf score for chapter 3.

From this, we could make some guesses as to what the chapters in this book are about. Perhaps milk was spilled on the couch and somone needed money to buy new furniture in chapter 2.


## Unique Word Counts
Sometimes, we may want to see how many words occur in a document or set of documents. We may also want to know how many times that word occurs in a document or set of documets. We will go through a brief example of how to do that now.

First, we will create a dataframe that has a column counting the total number of times each word occurs across the entire novel.

```{r warning=FALSE, message=FALSE}
library(data.table)
total_word_counts <- meta %>%
  unnest_tokens(word, text) %>%
  add_count(word)
```

Next, we use our dataframe, 'book_words' from the tf-idf example, which has a column counting the number of times each word occurs in a chapter.

```{r warning=FALSE, message=FALSE}
book_words <- meta %>%
  unnest_tokens(word, text) %>%
  add_count(chapter, word, sort = TRUE)

book_words <- distinct(book_words)
```

Joining the two dataframes together and changing column names appropriately:

```{r warning=FALSE, message=FALSE}
word_counts <- book_words %>%
  left_join(total_word_counts, by = c("word","chapter"))

word_counts <- distinct(word_counts)


setnames(word_counts, old = "n.x", new = "chapter_count")
setnames(word_counts, old = "n.y", new = "book_count")

word_counts %>%
  arrange(desc(chapter_count))
```

Now we can see unique word counts by chapter and across the entire book.

As you can see, there are many ways to give importance to words in text data. Word/phrase counts (term frequency) and tf-idf are two great places to start when analytzing text data.
