---
title: "Tokenizing with tidytext"
author: "Team Hanley NLP Working Group"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting the Data

We will be using the book Metamorphosis by Franz Kafka as our example data. 

```{r warning=FALSE, message=FALSE}

library(dplyr)
library(tidytext)
library(gutenbergr)
library(kableExtra)
library(tidyr)

```
## Tokenizing by n-grams

Tokenization is the process of breaking your text into pieces or tokens. As noted [here](https://nlp.stanford.edu/IR-book/html/htmledition/tokenization-1.html), a token is "an instance of a sequence of characters in some particular document that are grouped together as a useful semantic unit for processing." In practice, tokens often refer to words. 

### Unigrams

An n-gram is a contiguous sequence of n items from a sample of text. An n-gram of size 1 is called a unigram, size 2 is a bigram, and size 3 is a trigram. 

Here's an example of how to tokenize by unigrams. We'll use our text from Metamorphosis.

```{r, message=F, warning=F, error = F}

# downloading metamorphasis
meta = gutenberg_download("5200")

unigrams = meta %>% unnest_tokens(word, text, token = "ngrams", n = 1)

unigrams%>%
    # below is purely for visuzliations purposes
  slice(1:5)%>%
  kable()%>%
  kable_styling("striped")

```

### Bigrams

And here's an example of tokenizing by bigrams. Notice the only argument in `unnest_tokens` that needs to change from our unigram example is `n =`. (The first argument listed here is the column name, which can be whatever you want.)

```{r, message=F, warning=F, error = F}


bigrams = meta %>% unnest_tokens(bigram, text, token = "ngrams", n = 2)

# below is purely for visuzliations purposes
bigrams%>%
  slice(1:5)%>%
  kable()%>%
  kable_styling("striped")

```


## Removing Stop Words
Stop words are words that are typically more important to the grammer of a sentence rather than the meaning of a sentence. They are like helper words that help string together meaning without providing much meaning themselves. 

For humans, more words are needed but for NLP they add nothing. Let's look at that last sentence for example.

>For humans, more words are needed but for NLP they add nothing.

Now without the stop words!

>humans more words needed NLP add nothing.

As you can see, computer really only need cave man speak to get the semantic meaning of words.

### How do we know if something is a stop word?
As you might guess, what exactly constitutes a stop word is somewhat subjective. When thinking about defining a stop word, we need to think about generalities and scale. For example, "no" can be a really important word semantically, and for certain tasks, like sentiment analysis, [you may want to keep certain stop words](https://medium.com/@limavallantin/why-is-removing-stop-words-not-always-a-good-idea-c8d35bd77214), especially words like "no" or "not". Handling these words is called [negation handling](https://towardsdatascience.com/sentiment-analysis-in-r-good-vs-not-good-handling-negations-2404ec9ff2ae). But we will get into that when we cover sentiment analysis.   

Lets use our bigram example from before. We are going to be basing our methodogly for removing stop words off of [tidytextmining](https://www.tidytextmining.com/ngrams.html#counting-and-filtering-n-grams). Its pretty simple and somewhat scalable.

First split the ngrams into their own columns. In this case, since its bigrams, the columns will be "word1" and "word2"

```{r warning = FALSE, message = FALSE}

bigrams_separated <- bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

```

Next for each "word" column, filter out all the stop words

```{r warning = FALSE, message = FALSE}

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

```

Finally, concatinate all the columns back together with `unite()`

```{r warning = FALSE, message = FALSE}

bigrams_united <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ") 

bigrams_united%>%
  slice(1:5)%>%
  kable()%>%
  kable_styling("striped")


```

And there you have it! A clean dataframe of bigrams without stop words the `tidytext` way.

## Counting n-grams

While it may seem simple, counting frequency of n-grams (in our case bigrams) will tell us a lot about the content of the text. Let's look at the bigram frequency of the *Metamorphosis* with and without removing stop words:

**With Stop Words**
```{r, warning=FALSE, message=F}
bigrams%>%
  count(bigram, sort = T)%>%
  # below is purely for visuzliations purposes
  slice(1:5)%>%
  kable()%>%
  kable_styling("striped")
```

**Without Stop Words**
```{r, warning=FALSE, message=F}
bigrams_united%>%
  count(bigram, sort = T)%>%
  # below is purely for visuzliations purposes
  slice(1:5)%>%
  kable()%>%
  kable_styling("striped")
```

So as we can see, before we remove the stop words, we largely get unremarkable bigrams at the top of the list. In fact, I would guess that you would "of the" as one of the top bigrams in nearly any book or paper. 

Once we remove the stop words, we begin to generate content that is specific to our book. Interestingly, "earn money" is one of the top five bigrams in the book!


## Example: Trigrams for Heart of Darkness {.tabset .tabset-fade}

Overall, this tutorial is simply meant to acquaint you with the basic premises of preparing text data for analysis, specifically in the tidytext universe. 

Now, try and repeat the above steps for "Heart of Darknes" by Joseph Conrad. But this time, lets use trigrams instead of bigrams. **The Gutenberg id is 219**.

### Problem

We will be organizng the book "Heart of Darkness" into **trigrams**. You will need all the packages and skill sets described above to prepare "Heart of Darkness" for analysis. 

**Hints**
- `tidytext` is mostly scalable but will require some minor edits to handle trigrams
- Make sure you remove stop words from *all three* of the trigrams.

### Solution

**Step 1: Download the Book**
```{r, warning = F, message=F}

darkness = gutenberg_download("219")

```

**Step 2: Tokenize the Dataset**
```{r, warning = F, message=F}

darkness %>%
  unnest_tokens(trigram, text, token = "ngrams", n = 3)%>%
  # below is purely for visuzliations purposes
  slice(1:5)%>%
  kable()%>%
  kable_styling("striped")

```

**Step 3: Remove the Stopwords**
```{r, warning = F, message=F}

darkness %>%
  unnest_tokens(trigram, text, token = "ngrams", n = 3)%>%
  separate(trigram, c("word1", "word2", "word3"), sep = " ")%>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)%>%
  filter(!word3 %in% stop_words$word)%>%
  unite(trigram, word1, word2, word3, sep = " ")%>%
  # below is purely for visuzliations purposes
  slice(1:5)%>%
  kable()%>%
  kable_styling("striped")

```

## Conclusion

As you can see, handling text data with tidytext is very simple. The big advantage of using this package is that it works with other tidyverse functions, and this framework can even be used for some analytic tasks such as sentiment analysis and topic modeling. 