---
title: "Twitter Data with rtweet"
author: "Team Hanley NLP Working Group"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started

### Getting a Developer Account
caitlin -  youtube video

### Saving Credentials in a .gitignore
corydon

## rtweet

explantion on what rtweet is: corydon

### Retrieving Topics
caitlin

### User Data
caitlin

### Posting a tweet
Corydon


[How to create a Twitter Developer Account](https://www.extly.com/docs/autotweetng_joocial/tutorials/how-to-auto-post-from-joomla-to-twitter/apply-for-a-twitter-developer-account/#apply-for-a-developer-account)


### Saving Credentials in a .gitignore


## TwitteR

### Retrieving Topics

Michael Bunting

### User Tweets

Michael Bunting



## Example 1: For users 

## Example 2:

## Conclusion
```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
